---
title: "Dispersal"
author: "Jean"
date: "2023-11-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(dplyr)
library(magrittr)
library(tidyr)
library(ggplot2)
library(reshape2)
library(tinytable)
library(igraph)
```

## Parameters

```{r, Params}
# Initial stock
X1 = .9
X2 = .9
# Growth
alpha1 = .85
alpha2 = .85

# Relative fencing efficiencies
beta1 = .8
beta2 = .8
base = 0

#Scaling for dispersal
k = 10

# Discounting
delta = .99

# Econ
p = 10
cA = 2
cB = 9

```

## Function

```{r, Functions}

dispersal = function(i,
                     j,
                     fi,
                     fj,
                     k. = k,
                     base. = base){
  if(j == 2){
    y = exp(- (base. + fi + beta1*fj)/k.)
  }else if(j==1){
    y = exp( - (base.+ fj + beta2*fi)/k.)
  }
  return(y)
}

partial_dispersal = function(i,
                             j,
                             fi,
                             fj,
                             type,
                             k. = k){
  if(!(type %in% c('own', 'cross'))){
    stop("derivative type is not 'own' nor 'cross'" )
  }else if(type == 'own'){
    y = -1/k * dispersal(i,j,fi,fj,k.)
  }else if(type == 'cross'){
    if(j == 1){
      y = -beta2/k * dispersal(i,j,fi,fj,k.)
    }else if(j==2){
      y = - beta1/k * dispersal(i,j,fi,fj,k.)
    }
  }
  return(y)
}

growth = function(x,alpha){
  return(x^ alpha)
}

marginal_growth = function(x,alpha){
  return(alpha * x^(alpha-1))
}

```

## FOCs for optimal solution

```{r, FOC}

e1 = seq(0,1,0.01)

data = data.frame(e1)%>%
  mutate(growth = growth(e1, alpha1))%>%
  mutate(marginal_growth = marginal_growth(e1, alpha1))

data %>%
  ggplot(aes(x = e1, y = marginal_growth))+
  geom_point()

# Optimal ei

e1 = function(f1,f2){
  y = 1/(delta*alpha1)*((p-cA)/((1 - dispersal(1,2,f1,f2))* (p-cA) + (p-cB) * dispersal(2,1,f1,f2)))
  y = y^(1/(alpha1 - 1))
  z = min(y,X1)
  return(z)
} 


e2 = function(f1,f2){
  y = 1/(delta*alpha2)*((p-cB)/((1 - dispersal(2,1,f1,f2))*(p-cB) + dispersal(1,2,f1,f2)*(p-cA)))
  y = y^(1/(alpha2 - 1))
  z = min(y, X2)
  return(z)
}


# Optimality for fencing? 

bound = cB - cA

FOCA = function(f1,f2){
  y  = partial_dispersal(2,1,f1,f2,'cross') *  growth(e2(f1,f2), alpha2) - partial_dispersal(1,2,f1,f2, 'own') * growth(e1(f1,f2),alpha1)
  return(y)
}

FOCB = function(f1,f2){
  y  = partial_dispersal(2,1,f1,f2,'own') *  growth(e2(f1,f2), alpha2) - partial_dispersal(1,2,f1,f2,'cross')*growth(e1(f1,f2),alpha1)
  return(y)
}
```


## Analysis

```{r, optimal escapement}

input_ = expand.grid(seq(0,50), seq(0,50))
result_ = data.frame(matrix(ncol = 2, nrow = 0))

for(row in 1:nrow(input_)){
  condA = FOCA(input_[row,1], input_[row,2]) * bound
  condB = FOCB(input_[row,1], input_[row,2]) * bound
  result_[row,] = c(condA, condB)
}
colnames(result_) = c('condA', 'condB')
result_ = cbind(input_,result_)

result_ = result_ %>%
  pivot_longer(-c('Var1', 'Var2'), 
               names_to = 'names', 
               values_to = 'values')

result_ %>%
  filter(Var2 %in% c(0, 7,8, 20, 50))%>%
  filter(names %in% c("condA", "condB"))%>%
  filter(Var1 < 15)%>%
  mutate(Var2 = as.factor(Var2))%>%
  ggplot(aes( x = Var1, color = names, y = values))+
  geom_point()+
  geom_hline(aes(yintercept = 0))+
  facet_grid(~Var2)
```

Results seem to point to $f_1^B = 0$ and $f_1^A \to \infty$.

```{r, welfare}

welfare = function(f1,f2){
  y = (p - cA) * (X1 - e1(f1,f2)) + (p-cB)* (X2 - e2(f1,f2))
  z = (p - cA) * ((1 - dispersal(1,2,f1,f2)) * growth(e1(f1,f2), alpha1) +
                   dispersal(2,1,f1,f2) * growth(e2(f1,f2), alpha2)) + 
      (p-  cB) * ((1 - dispersal(2,1,f1,f2)) * growth(e2(f1,f2),alpha2) +
                  dispersal(1,2,f1,f2) * growth(e1(f1,f2), alpha1))
  return(y + delta * z)
}


model = function(f1, f2, scenario, rounder = 4){
   y = c(scenario, round(dispersal(1,2,f1,f2), rounder), round(dispersal(2,1,f1,f2),rounder), round(e1(f1,f2),rounder),
         round(e2(f1,f2),rounder), round(marginal_growth(e1(f1,f2), alpha1),rounder), 
         round(marginal_growth(e2(f1,f2), alpha2),rounder), 
         round(1/delta,rounder), 
         round(welfare(f1,f2),rounder))

  names(y) = c('Scenario', 'Dispersal from 1 to 2', 'Dispersal from 2 to 1', 'Optimal escapement in 1', 'Optimal escapement in 2', 'Marginal growth in 1', 'Marginal growth in 2', '1/delta', "Welfare")
  return(y)
}

```


## Analysis of results with no inflow reduction

Assume the inflow is not reduced, eg $\frac{\partial d_{ij}}{\partial f_j}=0$. Start from a fully fenced equilibrium, where everyone is fencing their land:

```{r, full fence}
# Relative fencing efficiencies
beta1 = 0
beta2 = 0
base = 0
# Map the connectivity flows: 
view_network = function(f1, f2, X1. = X1, X2. = X2){
  g = graph(edges = c(1,2, 2,1, 1,1, 2,2), directed = T)
  
  V(g)$name <- c("Node 1", "Node 2")
  
  # Set node sizes based on the input radii
  V(g)$size <- c(radius1, radius2)*20   # Multiplying by a factor for better visibility
  
  # Set edge (arrow) width
  E(g)$width <- c(dispersal(1,2,f1,f2),dispersal(2,1,f1,f2), 1 - dispersal(1,2,f1,f2),1 - dispersal(2,1,f1,f2))
  
  layout = layount_nicely(g)
  # Plot the graph
  plot(g, 
       layout=layout,  # Layout to position nodes in a circle
       vertex.shape="circle",  # Nodes are shaped as circles
       vertex.label=V(g)$name,  # Display node names
       edge.arrow.size = max(line_width)/5,
       edge.curved=0.5)  # Slightly curve the edges for better visibility
}



view_connectivity = function(f1,f2){
  d_ = matrix(nrow = 2, 
              ncol = 2)
  d_[1,1] = 1 - dispersal(1, 2, f1, f2)
  d_[1,2] = dispersal(1, 2, f1, f2)
  d_[2,1] = dispersal(2, 1, f1, f2)
  d_[2,2] = 1 - dispersal(2, 1, f1, f2)
  
  ggplot(melt(d_), aes(x = Var2, y = Var1, fill = value)) + 
    geom_tile() + 
    scale_fill_gradient(low = "gray95", high = "indianred") + # Define the gradient fill
    geom_text(aes(label = round(value,3)))+
    labs(x = "",y = "") +
    theme_bw()+
    theme(
    # Remove axis titles
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    # Remove axis text labels
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    # Remove axis ticks
    axis.ticks = element_blank(),
    # Remove panel background, border, and grid lines
    panel.background = element_blank(),
    panel.border = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # Set color white : 
    text = element_text(color = "white"),
    # Remove legend
    legend.position = "none"
    )
}
view_connectivity(100,100)

```

This is likely one of the Nash equilibria. In this setting, the model yields:

```{r, welfare with fences in A}
tt(as.data.frame(t(model(100,100, 'Nash equilibrium'))))

```
The key fact here is to notice that the Nash equilibrium solves the spatial externality, and that the marginal growth rate in each patch is equal to $1/\delta$. Welfare however, could be larger!


```{r, welfare analysis}
# Reduced fencing : 

f1 = .4
f2 = .8
view_connectivity(f1, f2)
dispersal(1,2,f1,f2)
dispersal(2,1, f1, f2)
d_ = matrix(nrow = 2, 
            ncol = 2)
d_[1,1] = 1 - dispersal(1,2,f1,f2)
d_[1,2] = dispersal(1,2,f1,f2)
d_[2,1] = dispersal(2,1,f1, f2)
d_[2,2] = 1 - dispersal(2,1,f1,f2)
print(d_)
```

In this situation: 

```{r, welfare source sink}
model(f1,f2, 'Low value is sink')
```

In this situation : 

* Optimal escapement in patch $B$ is constrained by initial size : no harvest, to allow for tomorrow to have a better money in $A$. 

* Harvest a lot today in $A$ to keep the resource. 

What one may be attracted to do in this case, given that fencing only reduces outflow but has no impact on inflow (`r print(beta1)`) is to retain more resource in $A$ (and possibly, to reverse the sink source dynamics) : 

# Investigation

```{r, checker}

input_ = expand.grid(seq(1,50), seq(1,50))
results = data.frame(matrix(ncol = 5))

for(i in 1:nrow(input_)){
  results[i,] = c(input_[i,1], input_[i,2], dispersal(1,2, input_[i,1],input_[i,2]), dispersal(2,1,input_[i,1],input_[i,2]), welfare(input_[i,1], input_[i,2]))
}

max(results$X5)

e1(0,0)
1/(delta*alpha1)*(p-cA)/(p-cB)
```